---
title: "Lending Club Loan Analysis R Notebook"
output: html_notebook
---

This project uses the lending club loan data set available on kaggle to explore the loan data and predict defaults using machine learning techniques. Our idea is to merge the dataset provided with external macro data to see if we can get additional insight and better predictions.

### Getting the Data

```{r}
loan_url<-"https://www.kaggle.com/wendykan/lending-club-loan-data/downloads/loan.csv.zip"
```
### Reading the dataset
```{r}
library(data.table)
library(tidyr)

loan<-fread(input = "lending-club-loan-data/loan.csv",header = T)

loan_lca<-fread(input = "Original Data/LoanStats3a_securev1.csv",header = T,skip = 1)
loan_lcb<-fread(input = "Original Data/LoanStats3b_securev1.csv",header = T,skip = 1)
loan_lcc<-fread(input = "Original Data/LoanStats3c_securev1.csv",header = T,skip = 1)
loan_lcd<-fread(input = "Original Data/LoanStats3d_securev1.csv",header = T,skip = 1)

loan_lc<-rbind(loan_lca,loan_lcb,loan_lcc,loan_lcd)
```
### Data View
```{r}
head(loan_lc,10)

```

### Data Summary
```{r}
str(loan_lc)
```
### Selecting Column Subset
```{r}
initial_cols<-c("id","application_type","loan_amnt","funded_amnt","term","int_rate","grade","sub_grade","emp_length","home_ownership","annual_inc","verification_status","annual_inc_joint","issue_d","loan_status","purpose","zip_code","addr_state","dti","dti_joint","earliest_cr_line","fico_range_low","fico_range_high","inq_last_6mths","open_acc","total_acc","pub_rec","revol_util","last_pymnt_d","next_pymnt_d","last_credit_pull_d","last_fico_range_high","last_fico_range_low","total_pymnt","total_pymnt_inv","total_rec_prncp","total_rec_int","total_rec_late_fee","recoveries","out_prncp","out_prncp_inv")

loan_lc_sub<-loan_lc[,initial_cols,with=FALSE]

numeric_cols<-c("loan_amnt","funded_amnt","int_rate","annual_inc","annual_inc_joint","dti","dti_joint","fico_range_low","fico_range_high","inq_last_6mths","last_fico_range_high","last_fico_range_low","open_acc","total_acc","revol_util","total_pymnt","total_pymnt_inv","total_rec_prncp","total_rec_int","total_rec_late_fee","recoveries","out_prncp","out_prncp_inv")

factor_cols<-c("id","application_type","term","grade","sub_grade","emp_length","home_ownership","verification_status","loan_status","purpose","zip_code","addr_state","pub_rec")

date_cols<-c("issue_d","earliest_cr_line")
## ,"last_pymnt_d","next_pymnt_d","last_credit_pull_d")

```
### Cleaning Data
```{r}
library(lubridate)
library(stringr)
library(stringi)


## Cleaning the interest rate column to remove % sign
loan_lc_sub$int_rate<-gsub("%","",x = loan_lc_sub$int_rate)

## Cleaning up term column to remove trailing spaces
loan_lc_sub$term<-str_trim(loan_lc_sub$term)

loan_lc_sub[,(numeric_cols):= lapply(.SD, as.numeric), .SDcols = numeric_cols]
loan_lc_sub[,(factor_cols):= lapply(.SD, as.factor), .SDcols = factor_cols]
```
### Clean up dates
```{r}
loan_lc_sub$issue_d<-as_date(ifelse(stri_length(loan_lc_sub$issue_d)==6,
                                    as.Date(gsub("^","01-",x=loan_lc_sub$issue_d),format="%d-%b-%y"),
                                    as.Date(gsub("^","01-",x=loan_lc_sub$issue_d),format="%d-%b-%Y") ))

loan_lc_sub$earliest_cr_line<-as_date(ifelse(stri_length(loan_lc_sub$earliest_cr_line)==6,
                                    as.Date(gsub("^","01-",x=loan_lc_sub$earliest_cr_line),
                                            format="%d-%b-%y",origin = "1900-01-01"),
                                    as.Date(gsub("^","01-",x=loan_lc_sub$earliest_cr_line),
                                            format="%d-%b-%Y",origin = "1900-01-01") ))

## to clean up older credit histories which are shown as future credit histories. about 412 records

loan_lc_sub[earliest_cr_line>"2015-01-01",earliest_cr_line:=earliest_cr_line-years(100)]

loan_lc_sub$last_pymnt_d<-as_date(ifelse(stri_length(loan_lc_sub$last_pymnt_d)==6,
                                    as.Date(gsub("^","01-",x=loan_lc_sub$last_pymnt_d),
                                            format="%d-%b-%y",origin = "1900-01-01"),
                                    as.Date(gsub("^","01-",x=loan_lc_sub$last_pymnt_d),
                                            format="%d-%b-%Y",origin = "1900-01-01") ))

summary(loan_lc_sub)
```
### Creating new attributes and prediction columns

```{r}
loan_test<-loan_lc_sub

loan_test$PERFORM<-as.factor(ifelse(loan_test$loan_status %in% c("Current","Fully Paid"),1,0))
loan_test[,RETURN := ((total_pymnt/funded_amnt)-1)/ifelse(loan_test$term %in% c("36 months"),3,6) ]
loan_test$AGE <- loan_test$last_pymnt_d-loan_test$issue_d

loan_defaults<-loan_test[PERFORM==0,.(id,loan_amnt,total_pymnt,term,int_rate,grade,emp_length,annual_inc,dti,
                                      issue_d,last_pymnt_d,purpose,addr_state,earliest_cr_line,fico_range_low,RETURN,AGE,PERFORM)]

loan_paid<-loan_test[PERFORM==1,.(id,loan_amnt,total_pymnt,term,int_rate,grade,emp_length,annual_inc,dti,
                                issue_d,last_pymnt_d,purpose,addr_state,earliest_cr_line,fico_range_low,RETURN,AGE,PERFORM)]

```

### Plotting differences between paid and default loans

```{r}
par(mfrow=c(2,1))
hist(as.numeric(loan_defaults$AGE))
hist(as.numeric(loan_paid$AGE))
```
- Number of loans alive as a proportion of loans alive at previous month

```{r}
par(mfrow=c(2,1))
hist(as.numeric(loan_defaults$RETURN))
hist(as.numeric(loan_paid$RETURN))
```

### Joining with unemployment data
```{r}
loans<-rbind(loan_defaults,loan_paid)
unemp.data<-read.csv("State Unemployment Data Monthly History.csv",header = T)
unemp.data$state<-as.factor(str_trim(unemp.data$state))
unemp.data$date<-as_date(unemp.data$date)
unemp.data<-as.data.table(unemp.data)
```
### Getting uneployment rate for each observation
```{r}

getUnempRate<-function(state_input,date_input){
    rate<-as.numeric(unemp.data[(str_trim(state)==str_trim(state_input))&(date==date_input),value])
    return(rate)
}

loans[unemp.data,UNEMP_RATE_ISSUE_DT:=i.value,on=c(addr_state="state",issue_d="date")]
loans[unemp.data,UNEMP_RATE_LASTPAY_DT:=i.value,on=c(addr_state="state",last_pymnt_d="date")]

```

### Marking Prepaid loans and maturity of loan
```{r}
loans[,MATURITY:=issue_d+months(36)][,PREPAID:=0]
loans[(PERFORM==1)&(year(last_pymnt_d)<2017),PREPAID:=ifelse(last_pymnt_d<MATURITY,1,0)]


```

```{r}
write.csv(x = loans,file = "LC Loan Data.csv",row.names = F)
```

