---
title: "Lending Club - Macro Effect"
output: html_notebook
---
 _Effect of Macro factors such as Unemployment Rate on Lending Club Loans_

### Summary

Six different regressions are performed to analyze the lending club loans. They are ordered as below

1. California Loans 2012-2015 Issuance
2. All State Loans 2012Q1 - 2013Q3 Issuance (using Unemployment Rate at Issuance)
3. All State Loans 2012Q1 - 2013Q3 Issuance (using Unemployment Rate Monthly)
4. All State Loans 2011 - 2015 Issuance (using Unemployment Rate Monthly)
5. All State Loans 2011 - 2015 Issuance (using Unemployment Rate at Issuance)
6. All State Loans 2011 - 2015 Issuance (using FICO scores + Unemp Issuance + Unemp Monthly)

In all the models above `Interest Rate` at issuance is included in the model

---

#### Load Libraries
```{r}
library(data.table)
library(tidyverse)
library(lubridate)
library(survival)
library(survminer)
```

---

### Analyze Lending Club loans 2012-2015 
**(California)**

Sample Data is for loans issues from 2012 through to 2015 onwards in the state of California with a 36 month term. A random sample of 1000 such loans is chosen for the analysis
```{r}
loan_sample<-read.csv("Sample Loans CA 36 2012_2015.csv", header=T)
head(loan_sample)
```

---

#### Selecting Data 
**(California)**

- Remove records for loan status after Oct 2016 (inclusive) to avoid dealing with Delinquent loans
- Normalize Unemployment Rate at begin date and interest rate columns by subtracting means
- Select the columns which are going to be used in the model only
- Factorize the age of the loan

```{r}
loan_sample_model<-as.tbl(loan_sample)%>%
    filter(as.Date(MONTH_BEGIN_PAY_DT,origin=Sys.Date()) < 
               as.Date("2016-10-01",origin=Sys.Date()))%>%
    select(id,int_rate,MONTH.AGE.BEGIN,DEFAULT,UNEMP.RT.BEGIN_DT)%>%
    mutate(int_rate = int_rate-mean(int_rate))%>%
    mutate(UNEMP.RT.BEGIN_DT = UNEMP.RT.BEGIN_DT-mean(UNEMP.RT.BEGIN_DT))%>%
    mutate(MONTH.AGE.BEGIN = as.factor(MONTH.AGE.BEGIN))

head(loan_sample_model)
```

---

#### Fit a Linear Model
**(California)**
```{r}
loan_linear_mdl<-lm(DEFAULT ~ int_rate + UNEMP.RT.BEGIN_DT+MONTH.AGE.BEGIN - 1,
                     data = loan_sample_model)

summary(loan_linear_mdl)

```

---

#### Baseline Hazard Rate for Loans 
**(California)**
```{r}
par(mar = c(3,3,4,2))
plot(x = loan_linear_mdl$xlevels$MONTH.AGE.BEGIN,
     y = loan_linear_mdl$coefficients[-(1:2)],
     type = 'h',
     lwd=5,
     col = 'blue',
     xlab = "Age of Loan (months)",
     ylab = "Default Rate",
     main = "California Lending Club Loans Sample (2012-2015)")
```


---

### Analyzing All States Sample for loans 
**(2012-2013Q3 Unemployment Issuance)**

Sample Data is for loans issues from 1-Jan-2012 through to 30-Sep-2013, in all the states in the US with a 36 month term. A random sample of 10,000 such loans is chosen for the analysis
```{r}
loan_sample_all_2012H1<-fread("Sample Loans 36 2012_2013.csv", header=T)
head(loan_sample_all_2012H1)
```

---

#### Selecting Data 
**(2012-2013Q3 Unemployment Issuance)**

- Remove records for loan status after Oct 2016 (inclusive) to avoid dealing with Delinquent loans
- Normalize Unemployment Rate at begin date and interest rate columns by subtracting means
- Select the columns which are going to be used in the model only
- Factorize the age of the loan

```{r}
loan_sample_all_model_2012H1<-as.tbl(loan_sample_all_2012H1)%>%
    filter(as.Date(MONTH_BEGIN_PAY_DT,origin=Sys.Date()) < 
               as.Date("2016-10-01",origin=Sys.Date()))%>%
    select(id,int_rate,MONTH.AGE.BEGIN,DEFAULT,UNEMP.RT_ISS)%>%
    mutate(int_rate = int_rate-mean(int_rate))%>%
    mutate(UNEMP.RT_ISS = UNEMP.RT_ISS-mean(UNEMP.RT_ISS))%>%
    mutate(MONTH.AGE.BEGIN = as.factor(MONTH.AGE.BEGIN))

head(loan_sample_all_model_2012H1)
```

---

#### Fit a Linear Model 
**(2012-2013Q3 Unemployment Issuance)**
```{r}
loan_linear_mdl_all_2012H1<-lm(DEFAULT ~ int_rate + UNEMP.RT_ISS+MONTH.AGE.BEGIN - 1,
                     data = loan_sample_all_model_2012H1)

summary(loan_linear_mdl_all_2012H1)

```

---

#### Baseline Hazard Rate for ALL States Loans 
**(2012-2013Q3 Unemployment Issuance)**
```{r}
par(mar = c(3,3,4,2))
plot(x = loan_linear_mdl_all_2012H1$xlevels$MONTH.AGE.BEGIN,
     y = loan_linear_mdl_all_2012H1$coefficients[-(1:2)],
     type = 'h',
     col = 'darkgreen',
     lwd =5,
     xlab = "Age of Loan (months)",
     ylab = "Default Rate",
     main = "Lending Club Loans Sample (2012-2013Q3)")
mtext("Controlled for Unemployment Rate at Issuance",col="red",side=3)
```


---

### Analyzing All States Sample for loans 
**(2012-2013Q3 Unemployment Monthly)**

Sample Data is for loans issues from 1-Jan-2012 through to 30-Sep-2013, in all the states in the US with a 36 month term. A random sample of 10,000 such loans is chosen for the analysis
```{r}
loan_sample_all_2012H2<-fread("Sample Loans 36 2012_2013.csv", header=T)
head(loan_sample_all_2012H2)
```

---

#### Selecting Data
**(2012-2013Q3 Unemployment Monthly)**

- Remove records for loan status after Oct 2016 (inclusive) to avoid dealing with Delinquent loans
- Normalize Unemployment Rate at begin date and interest rate columns by subtracting means
- Select the columns which are going to be used in the model only
- Factorize the age of the loan

```{r}
loan_sample_all_model_2012H2<-as.tbl(loan_sample_all_2012H2)%>%
    filter(as.Date(MONTH_BEGIN_PAY_DT,origin=Sys.Date()) < 
               as.Date("2016-10-01",origin=Sys.Date()))%>%
    select(id,int_rate,MONTH.AGE.BEGIN,DEFAULT,UNEMP.RT.BEGIN_DT)%>%
    mutate(int_rate = int_rate-mean(int_rate))%>%
    mutate(UNEMP.RT.BEGIN_DT = UNEMP.RT.BEGIN_DT-mean(UNEMP.RT.BEGIN_DT))%>%
    mutate(MONTH.AGE.BEGIN = as.factor(MONTH.AGE.BEGIN))

head(loan_sample_all_model_2012H2)
```

---

#### Fit a Linear Model 
**(2012-2013Q3 Unemployment Monthly)**
```{r}
loan_linear_mdl_all_2012H2<-lm(DEFAULT ~ int_rate + UNEMP.RT.BEGIN_DT + MONTH.AGE.BEGIN - 1,
                     data = loan_sample_all_model_2012H2)

summary(loan_linear_mdl_all_2012H2)

```

---

#### Baseline Hazard Rate for ALL States Loans 
**(2012-2013Q3 Unemployment Monthly)**
```{r}
par(mar = c(3,3,4,2))
plot(x = loan_linear_mdl_all_2012H2$xlevels$MONTH.AGE.BEGIN,
     y = loan_linear_mdl_all_2012H2$coefficients[-(1:2)],
     type = 'h',
     col = 'violet',
     lwd =5,
     xlab = "Age of Loan (months)",
     ylab = "Default Rate",
     main = "Lending Club Loans Sample (2012-2013Q3)")
mtext("Controlled for Unemployment Rate Monthly",col="red",side=3)
```

---

### Analyzing All States Sample 
**(2011-2015 Unemployment Monthly)**

Sample Data is for loans issues from 2011 through to 2015 onwards in all the states in the US with a 36 month term. A random sample of 100,000 such loans is chosen for the analysis
```{r}
loan_sample_all<-fread("Sample Loans 36 2011_2015.csv", header=T)
head(loan_sample_all)
```

---

#### Selecting Data
**(2011-2015 Unemployment Monthly)**

- Remove records for loan status after Oct 2016 (inclusive) to avoid dealing with Delinquent loans
- Normalize Unemployment Rate at begin date and interest rate columns by subtracting means
- Select the columns which are going to be used in the model only
- Factorize the age of the loan

```{r}
loan_sample_all_model<-as.tbl(loan_sample_all)%>%
    filter(as.Date(MONTH_BEGIN_PAY_DT,origin=Sys.Date()) < 
               as.Date("2016-10-01",origin=Sys.Date()))%>%
    select(id,int_rate,MONTH.AGE.BEGIN,DEFAULT,UNEMP.RT.BEGIN_DT)%>%
    mutate(int_rate = int_rate-mean(int_rate))%>%
    mutate(UNEMP.RT.BEGIN_DT = UNEMP.RT.BEGIN_DT-mean(UNEMP.RT.BEGIN_DT))%>%
    mutate(MONTH.AGE.BEGIN = as.factor(MONTH.AGE.BEGIN))

head(loan_sample_all_model)
```

---

#### Fit a Linear Model 
**(2011-2015 Unemployment Monthly)**
```{r}
loan_linear_mdl_all<-lm(DEFAULT ~ int_rate + UNEMP.RT.BEGIN_DT+MONTH.AGE.BEGIN - 1,
                     data = loan_sample_all_model)

summary(loan_linear_mdl_all)

```


---

#### Baseline Hazard Rate for ALL States Loans 
**(2011-2015 Unemployment Monthly)**
```{r}
par(mar=c(3,3,4,2))
plot(x = loan_linear_mdl_all$xlevels$MONTH.AGE.BEGIN,
     y = loan_linear_mdl_all$coefficients[-(1:2)],
     type = 'h',
     col = 'red',
     lwd=5,
     xlab = "Age of Loan (months)",
     ylab = "Default Rate",
     main = "Lending Club Loans Sample (2011-2015)")
mtext("Controlled for Unemployment Rate Monthly",col="red",side=3)
```

---

### Analyzing and Selecting Data ALL states Loans 
**(2011-2015 Unemployment Issue Date)** 

```{r}
loan_sample_all_model_uis<-as.tbl(loan_sample_all)%>%
    filter(as.Date(MONTH_BEGIN_PAY_DT,origin=Sys.Date()) < 
               as.Date("2016-10-01",origin=Sys.Date()))%>%
    select(id,int_rate,MONTH.AGE.BEGIN,DEFAULT,UNEMP.RT_ISS,UNEMP.RT.BEGIN_DT)%>%
    mutate(int_rate = int_rate-mean(int_rate))%>%
    mutate(UNEMP.RT.BEGIN_DT = UNEMP.RT.BEGIN_DT-mean(UNEMP.RT.BEGIN_DT))%>%
    mutate(UNEMP.RT_ISS = UNEMP.RT_ISS-mean(UNEMP.RT_ISS))%>%
    mutate(MONTH.AGE.BEGIN = as.factor(MONTH.AGE.BEGIN))

head(loan_sample_all_model_uis)
```

---

#### Fit a Linear Model 
**(2011-2015 Unemployment Issue Date)**
```{r}
loan_linear_mdl_all_uis<-lm(DEFAULT ~ int_rate + UNEMP.RT_ISS + MONTH.AGE.BEGIN - 1,
                     data = loan_sample_all_model_uis)

summary(loan_linear_mdl_all_uis)

```


---

#### Baseline Hazard Rate for ALL States Loans 
**(2011-2015 Unemployment Issue Date)**
```{r}
par(mar = c(3,3,4,2))
plot(x = loan_linear_mdl_all_uis$xlevels$MONTH.AGE.BEGIN,
     y = loan_linear_mdl_all_uis$coefficients[-(1:2)],
     type = 'h',
     col = 'skyblue',
     lwd =5,
     xlab = "Age of Loan (months)",
     ylab = "Default Rate",
     main = "Lending Club Loans Sample (2011-2015)" )
mtext("Controlled for Unemployment Rate at Issuance",col="red",side=3)
```

---

### Analyzing loans (2011-2015) 
**With FICO Scores**

Sample Data is for loans issues from 2011 through to 2015, in all the states in the US with a 36 month term. A random sample of 100,000 such loans is chosen for the analysis
```{r}
loan_sample_all<-fread("Sample Loans 36 2011_2015.csv", header=T)
head(loan_sample_all)
```


---

#### Selecting Data
**With FICO Scores**

- Remove records for loan status after Oct 2016 (inclusive) to avoid dealing with Delinquent loans
- Normalize Unemployment Rate at begin date and interest rate columns by subtracting means
- Select the columns which are going to be used in the model only
- Factorize the age of the loan

```{r}
loan_sample_all_model_fico<-as.tbl(loan_sample_all)%>%
    filter(as.Date(MONTH_BEGIN_PAY_DT,origin=Sys.Date()) < 
               as.Date("2016-10-01",origin=Sys.Date()))%>%
    select(id,int_rate,fico_range_low,MONTH.AGE.BEGIN,DEFAULT,UNEMP.RT_ISS,UNEMP.RT.BEGIN_DT)%>%
    mutate(int_rate = int_rate-mean(int_rate))%>%
    mutate(UNEMP.RT.BEGIN_DT = UNEMP.RT.BEGIN_DT-mean(UNEMP.RT.BEGIN_DT))%>%
    mutate(UNEMP.RT_ISS = UNEMP.RT_ISS-mean(UNEMP.RT_ISS))%>%
    mutate(fico_range_low = fico_range_low-mean(fico_range_low))%>%
    mutate(MONTH.AGE.BEGIN = as.factor(MONTH.AGE.BEGIN))

head(loan_sample_all_model_fico)
```

---

#### Fit a Linear Model
**With FICO Scores**
```{r}
loan_linear_mdl_all_fico<-lm(DEFAULT ~ int_rate + fico_range_low + UNEMP.RT_ISS + 
                                 UNEMP.RT.BEGIN_DT + MONTH.AGE.BEGIN - 1,
                     data = loan_sample_all_model_fico)

summary(loan_linear_mdl_all_fico)

```

---

#### Baseline Hazard Rate for ALL States Loans
**With FICO Scores**
```{r}
par(mar = c(3,3,4,2),bg="wheat" )
plot(x = loan_linear_mdl_all_fico$xlevels$MONTH.AGE.BEGIN,
     y = loan_linear_mdl_all_fico$coefficients[-(1:4)],
     type = 'h',
     col = "brown",
     lwd =5,
     xlab = "Age of Loan (months)",
     ylab = "Default Rate",
     main = "Lending Club Loans Sample (2011-2015)" )
mtext("Controlled for FICO",col="red",side=3)
```

---

